On the presentation of orthogonal groups over polynomial rings

### 2019-06-13 dl-paper.tex
Many small fixes

\subsection{Steinberg groups}
%TODO: Describe what root systems are sympelctic and which are not
Let $\Phi$ be a reduced and irreducible root system of rank $\geq 2$ and $R$ be a commutative ring with $1$. Recall that in this case the \emph{Steinberg group} $\St(\Phi, R)$ can be defined by means of generators $x_{\alpha}(\xi)$ where $\xi\in R, \alpha\in\Phi$ and relations:
\begin{align}
& \phantom{[}
x_\alpha(s) x_\alpha(t) = x_\alpha(s+t),\ \alpha\in\Phi,\ s,t\in R; \label{rel:add}\\
& [x_\alpha(s), x_\beta(t)] = \prod\limits_{i,j\in\mathbb{N}}
 x_{i\alpha + j\beta}\left(N_{\alpha\beta ij}\, s^i t^j\right),\quad \alpha,\beta\in\Phi,\ \alpha\neq\pm\beta,\ s,t\in R. \label{rel:CCF}
\end{align}
The indices $i$, $j$ appearing in the right-hand side of the above relation range over
all positive natural numbers such that $i\alpha + j\beta\in\Phi$.
The structure constants $N_{\alpha \beta i j}=\pm 1,2,3$ appearing in \eqref{rel:CCF} depend only on $\Phi$ and can be computed explicitly.

@@

Recall that for $\alpha\in\Phi$ and $s \in R^*$ the semisimple root elements $h_\alpha(s)$ are defined as follows: \[h_\alpha(s)=w_\alpha(s) \cdot w_\alpha(-1), \text{ where } w_\alpha(s) = x_\alpha(s) \cdot x_{-\alpha}(-s^{-1}) \cdot x_\alpha(s).\]

@@

For our computations we will need two explicit elements of $\K_2(\Phi, R)$. Recall that Steinberg symbols are defined for any $s, t \in R^*$ by the identity
\begin{equation} \label{eq:steinberg} \{ s, t \}_\alpha = h_\alpha(s) \cdot h_\alpha(t) \cdot h_\alpha(st)^{-1}. \end{equation}

Our convention for Dennis--Stein symbols follows~\cite{DS73} (and {\it not} III.5.11 of ~\cite{Kbook}).
Recall that Dennis--Stein symbols are defined for arbitrary $r, s\in R$ satisfying $1 + rs \in R^*$ as follows:
\begin{equation} \label{eq:dennis-stein}  \langle r,s \rangle _ \alpha = x_{-\alpha}\left(\tfrac{- s}{1 + rs}\right) \cdot x_{\alpha}(r) \cdot x_{-\alpha}(s) \cdot x_{\alpha}\left(\tfrac{- r}{1+rs}\right) \cdot h_{\alpha}(1 + rs)^{-1}. \end{equation}

Dennis--Stein symbols and Steinberg symbols are related to each other via the following identities:
\begin{equation} \label{DS-S-relationship} \langle a, b \rangle_\alpha = \{-a, 1+ab\}_\alpha\text{ for } a, 1+ab\in R^*,\ \
 \{ s, t \}_\alpha = \left\langle -s, \tfrac{1 - t}{s} \right\rangle_\alpha\text{ for } s, t\in R^*. \end{equation}

@@

\subsection{Unstable Quillen K-groups} \label{sec:quillen}
Throughout this section $G, P, S$ denote three group-valued functors on the category of commutative rings satisfying the following reqirements:
\begin{enumerate}
 \item \label{req:left-exact} $G$ is left exact;
 \item \label{req:subfunc} $P$ is a subfunctor of $G$ such that for all $R$ the group $P(R)$ is a perfect normal subgroup of $G(R)$;
 \item \label{req:uce} There is a natural transformation $\varphi \colon S \to P$ such that $ \varphi \colon S(R) \to P(R)$ is a universal central extension for all $R$;
 \item \label{req:coeq} $S$ preserves coequalizers of kernel pairs.
\end{enumerate}

@@

%TODO: Give the definition of $D_{R, I}$ and of the category of Pairs
\begin{prop}\label{characterization}
\begin{enumerate}
 \item There are natural isomorphisms
  \[ \K_2(G, P, R) \cong \Ker(S(R) \to G(R)), \ \K_3(G, P, R) \cong \HH_3(S(R), \ZZ). \]
 \item The map $\mu \colon S(R, I) \to S(R)$ is the universal relative central extension of 
    the map $S(R) \to S(R/I)$. The kernel of $\mu$ is naturally isomorphic to the third relative homology group $\HH_3(S(R), S(R/I), \ZZ)$.
 \item There is a natural exact sequence:
 \[ \xymatrix{\K_3(G, P, R) \ar[r] & \K_3(G, P, R/I) \ar@{->>}[r] & \Ker(\mu).} \]
 \end{enumerate}   
\end{prop}
\begin{proof}
 The first claim follows from Exercises~1.8--1.9 of \cite[Chapter~IV]{Kbook} (here we use~\eqref{req:subfunc}--\eqref{req:uce}). %provided one verifies that all the intermediate isomorphisms are natural in $R$.

@@

Let us mention the example of a triple $(G,P,S)$ satisfying the above conditions.
\begin{example} \label{ex:orth}
 For $n \geq 5$ take the orthogonal group functor $\mathrm{O}_{2n}(-)$ and its elementary subfunctor $\mathrm{EO}_{2n}(-)$ as functors $G_n$ and $P_n$, respectively.
 Now set $S_n = \St(\rD_n, -)$.
 From [Stein's paper] and [Lemma on centrality] it follows that $S_n$ and $P_n$ satisfy~\eqref{req:uce}.
 The other requirements can be verified in a straightforward manner.
\end{example}

In the above definition we could have replaced the group $\mathrm{O}_{2n}$ with a different isogeneous form of the orthogonal group, e.\,g. 
 we could have chosen $G_n = \mathrm{SO}_{2n}$ or $G_n = \mathrm{Spin}_{2n}$ and then have defined $P_n$ as its elementary subfunctor.
Of course, this would change the functors $\K_i(G_n, P_n, -)$ for $i=1,2$ however, as the proof of~\cref{characterization} suggests %TODO: Does it really change K_2?
 it would not change $\K_i(G_n, P_n, -)$ for $i\geq 3$.

Now let $G_n, P_n$ be again as in~\cref{ex:orth}. We write $KO_i(n, R)$ as a shorthand for $\K_i(G_n, P_n, R)$.
Let us formulate the stability theorem for orthogonal Quillen K-groups (see~\cite[Theorem~9.4]{Pa89}).
\begin{theorem}[Panin] \label{lem:Panin-stability}
 Let $R$ be either a field, principal ideal domain or a Dedekind domain. Set $a = 1,2$ or $3$ in each of these respective cases.
 Then the canonical map $KO_i(2n, R) \to KO_i(2(n+1), R)$ is an epimorphism for $n \geq b$ 
 and an isomorphism for $\ell \geq b + 1$, where $b = \mathrm{max}(2i, a+i-1)$. \end{theorem}

@@

In the sequel we will need the following result.
\begin{theorem}[Bass Fundamental Theorem]\label{bass-ft} Suppose that $2 \in R^*$, then for any $i\geq 1$, $k\in \ZZ/4\ZZ$ there is a natural split exact sequence of abelian groups
 \[ \xymatrix{ 0 \ar[r] & \GW_i^{[k]}(R) \ar[r] & \GW_i^{[k]}(R[X, X^{-1}]) \ar[r]  & \GW_{i-1}^{[k-1]}(R) \ar[r] & 0.} \] \end{theorem}
\begin{proof} See~\cite[Corollary~5.3]{Ho05}. 
%This is a special case of~\cite[Theorem~9.13]{Sch16} and the remark that follows it. 
\end{proof}

@@

If we denote the kernel of the natural transformation $\mu$ from the definition of $\St(\Phi, R, I)$ by $C(\Phi, R, I)$ we obtain a natural exact sequence:
\begin{equation} \label{relSteinberg-def}
 \xymatrix{ 1 \ar[r] & C(\Phi, R, I) \ar[r]^{j} & \St(\Phi, R, I) \ar[r]^{\mu} & \St(\Phi, R) \ar[r]^{\pi} & \St(\Phi, R/I) \ar[r] & 1.}
\end{equation}
The group $C(\Phi, R, I)$ is trivial when $I$ is a splitting ideal of $R$, i.\,e. when the canonical map $R \to R/I$ splits.
This immediately follows e.\,g. from the exact sequence of~\cref{characterization}, it is also possible to give a direct proof of this fact, see \cite[Lemma~8]{S15}.
The following result plays a key role in the proof of~\cref{thm41}.
\begin{lemma} \label{lem:prop41}
In the notation of~\cref{lem:lemma32} assume additionally that the residue field $k$ is of characteristic $\neq 2$.
Then the canonical map $f\colon C(\rD_\ell, B, I) \to C(\rD_\ell, R, I)$ is surjective for $\ell \geq 7$.
\end{lemma}
\begin{proof}
  Consider the following commutative diagram with rows obtained from~\cref{characterization}:
\begin{equation*}\xymatrix{
 KO_3(2\ell, B) \ar[r] \ar[d] & KO_3(2\ell, k[X]) \ar[d]_{f'} \ar@{->>}[r] & \ar[d]_{f} C(\rD_\ell, B, I) \\
 KO_3(2\ell, R) \ar[r]        & KO_3(2\ell, k[X, X^{-1}]) \ar@{->>}[r]        & C(\rD_\ell, R, I).}\end{equation*}
In view of~\cref{lem:Panin-stability} the central vertical map $f'$ is isomorphic to the canonical map $\GW_3^{[0]}(k[X]) \to \GW_3^{[0]}(k[X, X^{-1}])$.
Invoking~\cref{bass-ft} we obtain that \[\GW_3^{[0]}(k[X, X^{-1}]) \cong \GW_3^{[0]}(k) \oplus \GW_2^{[3]}(k).\]
By \cite[Lemma~2.2]{FRS12} the group $\GW_2^{[3]}(k)$ is trivial, therefore $f'$ (and hence $f$) are surjective.
\end{proof}

\subsection{Some elementary lemmas}
\begin{comment}
Hall-Witt identity
\[ [[ y^{-1}, x], z] ^ {y^{-1}}  [[ z^{-1}, y], x] ^ {z^{-1}}  [[ x^{-1}, z], y] ^ {x^{-1}} = 1 \]
\[ [[ y^{-1}, x], z]  \cdot [[ z^{-1}, y], x] ^ {z^{-1}y} \cdot  [[ x^{-1}, z], y] ^ {x^{-1}y} = 1. \]
\[ [[ z^{-1}, y^{-1}], x] ^ {z^{-1}y^{-1}} \cdot  [[ x^{-1}, z], y^{-1}] ^ {x^{-1}y^{-1}} = [z, [ y, x]] . \]
\[  . \]
\end{comment}
Throughtout this section $A$ denotes an arbitary ring and $I, J$ two its ideals and $\Phi$ a simply laced root system.

The following lemma summarizes basic properties of structure constants which will be used in our computations below, often without further reference.
\begin{lemma} Suppose $(\alpha, \beta, \gamma)$ is a triple of roots forming a basis of root subsystem of type $\rA_3$
 (i.\,e. such that $\alpha+\beta, \beta + \gamma \in \Phi$ and $\alpha \perp \beta$), then holds:
\begin{equation} N_{\alpha, \beta} = -N_{\beta,\alpha} = - N_{-\alpha, -\beta} = N_{\beta, -\alpha-\beta} = N_{-\alpha-\beta, \alpha}. \end{equation}
\begin{equation} \label{eq:cocycle} N_{\beta,\gamma} \cdot N_{\alpha, \beta+\gamma} = N_{\alpha+\beta, \gamma} \cdot N_{\alpha, \beta}. \end{equation}
\end{lemma} \begin{proof} See~\cite[\S~14]{VP}. \end{proof}

@@

Now the same argument as in Lemma~\ref{welldef} allows to prove that $[s_{\alpha_{l-1}}(a),\,x_{\alpha_l}(b)]=1$ for $b\in R_e$, $e\leq0$, what finishes the proof of R3.2$\!\,_{n+1}$.
\end{comment}
\subsection{Computation of the kernel of the map of evaluation at \texorpdfstring{$0$}{0}.}
For a pair $(R, I)$ we denote by $\overline{\St}(\Phi, R, I)$ the image of the map $\mu$ from~\eqref{relSteinberg-def}, in particular, it is a normal subgroup of $\St(\Phi, R)$.
We also denote by $\St(\Phi, I)$ the subgroup of $\St(\Phi, R)$ generated as a group by root unipotents of level $I$.
It is clear that $\overline{\St}(\Phi, R, I)$ contains $\St(\Phi, I)$ and, in fact, is its normal closure.
We also denote by $\overline{H}(\Phi, R, I)$ the subgroup of $\overline{\St}(\Phi, R, I)$ generated by the semisimple root elements $h_\alpha(u)$ and symbols $\{u, v\}$, $u \in (1+I)^*$, $v \in R^*$, $\alpha\in \Phi$.

@@

\begin{rem}\label{Z-DS} Notice that our assumptions on the rank of $\Phi$ guarantee that $Z_0(\alpha)$ is nonempty, hence the group $Z_\alpha(A, M)$ contains relative Dennis--Stein symbols $\langle a, m \rangle$ for $a\in A$, $m\in M$.
By~\eqref{DS-S-relationship} relative Steinberg symbols $\{a, 1+m\}$ are also contained in $Z_\alpha(A, M)$ for all $a\in A^*$, $m\in M$. \end{rem}

@@

\subsection{The proof of the injectivity theorem}
Throughout this subsection $\Phi$ denotes arbitrary irreducible simply-laced root system of rank $\geq 3$, unless stated otherwise.
We denote by $B$ the ring $M[X] + A[X^{-1}]$ considered as a graded subring of $A[X, X^{-1}]$.
For shortness we also write $G_M^0$ instead of $\overline{\St}(\Phi, A, M)$.
\[ \xymatrix{ G_M^0 \ar@{^{(}->}[r]^{i_+} \ar@{^{(}->}[d]^{i_-} & \overline{\St}(\Phi, A[X], M[X]) \ar^{j_+^M}[d] \ar@{^{(}->}[r] & \St(\Phi, A[X]) \ar[dl]^{j_+} \\
              \St(\Phi, A[X\inv]) \ar[r]^{j_-} & \St(\Phi, A[X, X^{-1}]) &  } \] 
Denote by $\overline{G}^{\geq 0}_M$ the image of the map $j_+^M$. 
Denote by $V$ the subset of $\St(\Phi, A[X, X^{-1}]) \times \St(\Phi, A[X\inv]) \times (1 + M)^*$ consisting of those triples $(g, h, u)$ for which 
$p(g, h, u) := g \cdot j_-(h) \cdot \{ X, u \}$ belongs to $\overline{G}_M^{\geq 0}$.
We let $h_0 \in G_M^0$ act on $V$ on the right by $(g, h, u) \cdot h_0 = (g, h \cdot i_-(h_0), u)$.
It is clear that this action is well defined.
Denote by $\overline{V}$ the set of orbits of the action of $G_M^0$ on $V$.
Whenever $(g, h_1, u)$ and $(g, h_2, u)$ lie in the same orbit we use the notation $(g, h_1, u) \sim (g, h_2, u)$ and say that these triples are congruent.

>>>>

\subsection{Steinberg groups}
%TODO: Describe what root systems are sympelctic and which are not
Let $\Phi$ be a reduced and irreducible root system of rank $\geq 2$ and $R$ be a commutative ring with $1$. Recall that in this case the \emph{Steinberg group} $\St(\Phi, R)$ can be defined by means of generators $x_{\alpha}(\xi)$ where $\xi\in R, \alpha\in\Phi$ and relations:
\begin{align}
& \phantom{[}
x_\alpha(s) x_\alpha(t) = x_\alpha(s+t),\ \alpha\in\Phi,\ s,t\in R; \label{rel:add}\\
& [x_\alpha(s), x_\beta(t)] = \prod\limits_{i,j\in\mathbb{N}}
 x_{i\alpha + j\beta}\left(N_{\alpha\beta ij}\, s^i t^j\right),\quad \alpha,\beta\in\Phi,\ \alpha\neq\pm\beta,\ s,t\in R. \label{rel:CCF}
\end{align}
The indices $i$, $j$ appearing in the right-hand side of the above relation range over
all positive natural numbers such that $i\alpha + j\beta\in\Phi$.
The structure constants $N_{\alpha \beta i j}=\pm 1,\pm 2,\pm 3$ appearing in \eqref{rel:CCF} depend only on $\Phi$ and {\bf TODO}:
% в литературе описываются способы выбрать эти константы ссылка на третий взгляд.

@@

Recall that for $\alpha\in\Phi$ and $s \in R^\times$ the semisimple root elements $h_\alpha(s)$ are defined as follows: 
\[h_\alpha(s)=w_\alpha(s) \cdot w_\alpha(-1), \text{ where } w_\alpha(s) = x_\alpha(s) \cdot x_{-\alpha}(-s^{-1}) \cdot x_\alpha(s).\]

@@

In our computations we use two explicit families of elements of $\K_2(\Phi, R)$: Steinberg and Dennis--Stein symbols.
Recall that Steinberg symbols are defined for any $s, t \in R^\times$ by the identity
\begin{equation} \label{eq:steinberg} \{ s, t \}_\alpha = h_\alpha(s) \cdot h_\alpha(t) \cdot h_\alpha(st)^{-1}. \end{equation}

Our convention for Dennis--Stein symbols follows~\cite{DS73} (and {\it not} III.5.11 of ~\cite{Kbook}).
Recall that Dennis--Stein symbols are defined for arbitrary $r, s\in R$ satisfying $1 + rs \in R^\times$ as follows:
\begin{equation} \label{eq:dennis-stein}  \langle r,s \rangle _ \alpha = x_{-\alpha}\left(\tfrac{- s}{1 + rs}\right) \cdot x_{\alpha}(r) \cdot x_{-\alpha}(s) \cdot x_{\alpha}\left(\tfrac{- r}{1+rs}\right) \cdot h_{\alpha}(1 + rs)^{-1}. \end{equation}

Dennis--Stein symbols and Steinberg symbols are related to each other via the following identities:
\begin{equation} \label{DS-S-relationship} \langle a, b \rangle_\alpha = \{-a, 1+ab\}_\alpha\text{ for } a, 1+ab\in R^\times,\ \
 \{ s, t \}_\alpha = \left\langle -s, \tfrac{1 - t}{s} \right\rangle_\alpha\text{ for } s, t\in R^\times. \end{equation}

@@

\subsection{Unstable Quillen K-groups} \label{sec:quillen}
Throughout this section $G, P, S$ denote three group-valued functors on the category of commutative rings satisfying the following assumptions:
\begin{enumerate}
 \item \label{req:left-exact} $G$ is left exact;
 \item \label{req:subfunc} $P$ is a subfunctor of $G$ such that for all $R$ the group $P(R)$ is a perfect normal subgroup of $G(R)$;
 \item \label{req:uce} There is a natural transformation $\varphi \colon S \to P$ such that $ \varphi \colon S(R) \to P(R)$ is a universal central extension for all $R$;
 \item \label{req:coeq} $S$ preserves coequalizers of kernel pairs.
\end{enumerate}

@@

%TODO: Give the definition of $D_{R, I}$ and of the category of Pairs
\begin{prop}\label{characterization}
\begin{enumerate}
 \item There are natural isomorphisms
  \[ \K_2(G, P, R) \cong \Ker(S(R) \to G(R)), \ \K_3(G, P, R) \cong \HH_3(S(R), \ZZ). \]
 \item The map $\mu \colon S(R, I) \to S(R)$ is the universal relative central extension of the map $S(R) \to S(R/I)$.
       Denote the kernel of $\mu$ by $C(R, I)$.
       Then $C(R, I)$ is naturally isomorphic to the third relative homology group $\HH_3(S(R), S(R/I), \ZZ)$.
 \item There is a natural exact sequence:
 \[ \xymatrix{\K_3(G, P, R) \ar[r] & \K_3(G, P, R/I) \ar@{->>}[r] & C(R, I).} \]
 \end{enumerate}   
\end{prop}
\begin{proof}
 The first claim follows from \cite[Chapter~IV]{Kbook} (cf. Exercises~1.8--1.9 ibid, here we use assumptions~\eqref{req:subfunc}--\eqref{req:uce}). %provided one verifies that all the intermediate isomorphisms are natural in $R$.

@@

The principal example of a triple $(G,P,S)$ satisfying the above conditions is as follows.
For $n \geq 5$ take the orthogonal group functor $\mathrm{O}_{2n}(-)$ and its elementary subfunctor $\mathrm{EO}_{2n}(-)$ as functors $G_n$ and $P_n$, respectively. Now set $S_n = \St(\rD_n, -)$. From [Stein's paper] and [Lemma on centrality] it follows that $S_n$ and $P_n$ satisfy~\eqref{req:uce}.
 The other requirements can be verified in a straightforward manner. We write $KO_i(n, R)$ as a shorthand for $\K_i(G_n, P_n, R)$.

In the above definition we could have replaced the group $\mathrm{O}_{2n}$ with a different isogeneous form of the orthogonal group, e.\,g. 
 we could have chosen $G_n = \mathrm{SO}_{2n}$ or $G_n = \mathrm{Spin}_{2n}$ and then have defined $P_n$ as its elementary subfunctor.
Of course, this would change the functors $\K_i(G_n, P_n, -)$ for $i=1,2$ however, as the proof of~\cref{characterization} suggests
 it would not change $\K_3(G_n, P_n, -)$.

Let us formulate the stability theorem for orthogonal Quillen K-groups (see~\cite[Theorem~9.4]{Pa89}).
\begin{theorem}[Panin] \label{lem:Panin-stability}
 Let $R$ be either a field, principal ideal domain or a Dedekind domain. Set $a = 1,2$ or $3$ in each of these respective cases.
 Then the canonical map $KO_i(2n, R) \to KO_i(2(n+1), R)$ is an epimorphism for $n \geq b$ 
 and an isomorphism for $\ell \geq b + 1$, where $b = \mathrm{max}(2i, a+i-1)$. \end{theorem}

@@

In the sequel we will need the following result.
\begin{theorem}[Bass Fundamental Theorem]\label{bass-ft} Suppose that $2 \in R^*$, then for any $i\geq 1$, $k\in \ZZ/4\ZZ$ there is a natural split exact sequence of abelian groups
 \[ \xymatrix{ 0 \ar[r] & \GW_i^{[k]}(R) \ar[r] & \GW_i^{[k]}(R[X, X^{-1}]) \ar[r]  & \GW_{i-1}^{[k-1]}(R) \ar[r] & 0.} \] \end{theorem}
\begin{proof} This is a special case of~\cite[Theorem~9.13]{Sch16} and the remark that follows it. \end{proof}

@@

If we denote the kernel of the natural transformation $\mu$ from the definition of $\St(\Phi, R, I)$ by $C(\Phi, R, I)$ we obtain a natural exact sequence:
\begin{equation} \label{relSteinberg-def}
 \xymatrix{ 1 \ar[r] & C(\Phi, R, I) \ar[r]^{j} & \St(\Phi, R, I) \ar[r]^{\mu} & \St(\Phi, R) \ar[r]^{\pi} & \St(\Phi, R/I) \ar[r] & 1.}
\end{equation}
The group $C(\Phi, R, I)$ is trivial when $I$ is a splitting ideal of $R$, i.\,e. when the canonical map $R \to R/I$ splits, see \cite[Lemma~8]{S15}.
The following result plays a key role in the proof of~\cref{thm41}.
\begin{lemma} \label{lem:prop41}
In the notation of~\cref{lem:lemma32} assume additionally that the residue field $k$ is of characteristic $\neq 2$.
Then the canonical map $f\colon C(\rD_\ell, B, I) \to C(\rD_\ell, R, I)$ is surjective for $\ell \geq 7$.
\end{lemma}
\begin{proof}
  Consider the following commutative diagram with rows obtained from~\cref{characterization}:
\begin{equation*}\xymatrix{
 KO_3(2\ell, B) \ar[r] \ar[d] & KO_3(2\ell, k[X]) \ar[d]_{f'} \ar@{->>}[r] & \ar[d]_{f} C(\rD_\ell, B, I) \\
 KO_3(2\ell, R) \ar[r]        & KO_3(2\ell, k[X, X^{-1}]) \ar@{->>}[r]        & C(\rD_\ell, R, I).}\end{equation*}
In view of~\cref{lem:Panin-stability} the central vertical map $f'$ is precisely the canonical map $\GW_3^{[0]}(k[X]) \to \GW_3^{[0]}(k[X, X^{-1}])$.
Invoking~\cref{bass-ft} (or~\cite[Corollary~5.3]{Ho05}) we obtain that \[\GW_3^{[0]}(k[X, X^{-1}]) \cong \GW_3^{[0]}(k) \oplus \GW_2^{[3]}(k).\]
By \cite[Lemma~2.2]{FRS12} the group $\GW_2^{[3]}(k)$ is trivial, therefore $f'$ (and hence $f$) are surjective.
\end{proof}

\subsection{Some elementary lemmas}
\begin{comment}
Hall-Witt identity
\[ [[ y^{-1}, x], z] ^ {y^{-1}}  [[ z^{-1}, y], x] ^ {z^{-1}}  [[ x^{-1}, z], y] ^ {x^{-1}} = 1 \]
\[ [[ y^{-1}, x], z]  \cdot [[ z^{-1}, y], x] ^ {z^{-1}y} \cdot  [[ x^{-1}, z], y] ^ {x^{-1}y} = 1. \]
\[ [[ z^{-1}, y^{-1}], x] ^ {z^{-1}y^{-1}} \cdot  [[ x^{-1}, z], y^{-1}] ^ {x^{-1}y^{-1}} = [z, [ y, x]] . \]
\[  . \]
\end{comment}
Throughtout this section $A$ denotes an arbitary ring, and $I, J$ two its ideals, and $\Phi$ a simply laced root system.

The following lemma summarizes basic properties of structure constants which will be used in our computations below, often without further reference (cf.~\cite[\S~14]{VP}).
\begin{lemma} Suppose $(\alpha, \beta, \gamma)$ is a triple of roots forming a basis of root subsystem of type $\rA_3$
 (i.\,e. such that $\alpha+\beta, \beta + \gamma \in \Phi$ and $\alpha \perp \beta$), then holds:
\begin{equation} N_{\alpha, \beta} = -N_{\beta,\alpha} = - N_{-\alpha, -\beta} = N_{\beta, -\alpha-\beta} = N_{-\alpha-\beta, \alpha}. \end{equation}
\begin{equation} \label{eq:cocycle} N_{\beta,\gamma} \cdot N_{\alpha, \beta+\gamma} = N_{\alpha+\beta, \gamma} \cdot N_{\alpha, \beta}. \end{equation} \end{lemma}

@@

Now the same argument as in Lemma~\ref{welldef} allows to prove that $[s_{\alpha_{l-1}}(a),\,x_{\alpha_l}(b)]=1$ for $b\in R_e$, $e\leq0$, what finishes the proof of R3.2$\!\,_{n+1}$.
\end{comment}
\subsection{Computation of the kernel of the map of evaluation at \texorpdfstring{$0$}{0}.}
For a pair $(R, I)$ we denote by $\overline{\St}(\Phi, R, I)$ the image of the map $\mu$ from~\eqref{relSteinberg-def}, in particular, it is a normal subgroup of $\St(\Phi, R)$.
We also denote by $\St(\Phi, I)$ the subgroup of $\St(\Phi, R)$ generated as a group by root unipotents of level $I$.
It is clear that $\overline{\St}(\Phi, R, I)$ contains $\St(\Phi, I)$ and, in fact, is its normal closure.
We also denote by $\overline{H}(\Phi, R, I)$ the subgroup of $\overline{\St}(\Phi, R, I)$ generated by the semisimple root elements $h_\alpha(u)$ and symbols $\{u, v\}$, $u \in (1+I)^\times$, $v \in R^\times$, $\alpha\in \Phi$.

@@

\begin{rem}\label{Z-DS} Notice that our assumptions on the rank of $\Phi$ guarantee that $Z_0(\alpha)$ is nonempty, hence the group $Z_\alpha(A, M)$ contains relative Dennis--Stein symbols $\langle a, m \rangle$ for $a\in A$, $m\in M$.
By~\eqref{DS-S-relationship} relative Steinberg symbols $\{a, 1+m\}$ are also contained in $Z_\alpha(A, M)$ for all $a\in A^\times$, $m\in M$. \end{rem}

@@

\subsection{The proof of the injectivity theorem}
Throughout this subsection $\Phi$ denotes arbitrary irreducible simply-laced root system of rank $\geq 3$, unless stated otherwise.
We denote by $B$ the ring $M[X] + A[X^{-1}]$ considered as a graded subring of $A[X, X^{-1}]$.
For shortness we also write $G_M^0$ instead of $\overline{\St}(\Phi, A, M)$.
\[ \xymatrix{ G_M^0 \ar@{^{(}->}[r]^{i_+} \ar@{^{(}->}[d]^{i_-} & \overline{\St}(\Phi, A[X], M[X]) \ar^{j_+^M}[d] \ar@{^{(}->}[r] & \St(\Phi, A[X]) \ar[dl]^{j_+} \\
              \St(\Phi, A[X\inv]) \ar[r]^{j_-} & \St(\Phi, A[X, X^{-1}]) &  } \] 
Denote by $\overline{G}^{\geq 0}_M$ the image of the map $j_+^M$. 
Denote by $V$ the subset of $\St(\Phi, A[X, X^{-1}]) \times \St(\Phi, A[X\inv]) \times (1 + M)^\times$ consisting of those triples $(g, h, u)$ for which 
$p(g, h, u) := g \cdot j_-(h) \cdot \{ X, u \}$ belongs to $\overline{G}_M^{\geq 0}$.
We let $h_0 \in G_M^0$ act on $V$ on the right by $(g, h, u) \cdot h_0 = (g, h \cdot i_-(h_0), u)$.
It is clear that this action is well defined.
Denote by $\overline{V}$ the set of orbits of the action of $G_M^0$ on $V$.
Whenever $(g, h_1, u)$ and $(g, h_2, u)$ lie in the same orbit we use the notation $(g, h_1, u) \sim (g, h_2, u)$ and say that these triples are congruent.
